Bag of Words Semantic Object Classification Based on RGBD Images

### 2014-10-14 docs/piap/monografia/kkr13-eng.tex
po polsku

\title{3D Object Classification Based on RGBD Images}
\author{Adam Kosiorek}
\maketitle
Affiliation: Institute of Automation and Robotics, The Faculty of Mechatronics 
of Warsaw University of Technology \\
website: http://iair.mchtr.pw.edu.pl \\
Supervisor: Prof. dr hab. Barbara Siemiątkowska \\
Defense date: 11.02.2014

@@

 This paper introduces a Bag of Words (BoW) semantic object classification 
framework based on RGBD data registered by a Microsoft Kinect camera. We 
introduce a BoW image processing 
pipeline, describe it's steps in detail (region detection, feature 
extraction and vector quantization) and we outline the most successful 
algorithms. Algorithms with high performance potential were identified and 
evaluated. 
Salient region detection is done by SIFT, ISS3D and Uniform Sampling. 
Description is performed by FPFH, PFH and PFHRGB. Vector quantization was 
carried out by KMeans. An SVM with the RBF kernel trained with an 
All-vs-All method was used as a classifier. We achieved the accuracy of 65.22\% 
on 8 
categories of the Berkely 3D Object Dataset and 62.3\% on a 10 category dataset 
compiled by Zhang.

@@

@@

  It has never been easier to capture visual information. Popular storage 
services grow lighting fast due to terabytes of photographs and movies uploaded 
every day. The growth is so fast that hand-tagging and description, the 
traditional means of annotation, cease to suffice. They are ambiguous, 
emotional and rarely optimal. With no better solution at hand databases are 
becoming increasingly harder to browse. Similar issues occur in mobile 
robotics, where a robot has to understand it's enviornment if it is to interact 
with it. Both problems can be cast as a scene or object classification problem, 
one of the most popular computer vision issues in the last decade.

  This paper reviews the Bag of Words (or BoW for short) approach to image 
classification, evaluates available point cloud databases in terms of usability 
in scene classification studies and applies BoW model in the 3D domain. The 
rest of this work is organised as follows: Section 2 describes the Bag of Words 
technique in detail. Construction of an image model is discussed and the most 
popular algorithms are outlined and the problem of classification is 
introduced. Section 3 describes two renown RGBD image datasets suitable for 
evaluation of RGBD data-based object classifiers. Section 4 addresses the 
topics of experimental setup, conducted experiments and provides the results. 
Finally, Section 5 concludes this work.

@@

  The Bag of Words model originated from the natural language processing domain 
\cite{tsai2012bag}. It 
assumes that word occurrence in a text document carries semantic information. 
It is an intermediate representation (in a form of a hisogram)
used in natural language processing, information retrieval and general data 
mining.

  \begin{figure}[!ht]
  \centering
  \includegraphics[width=0.7\textwidth]{../figs/bow_example}
  \caption{Bag of Words histogram}
  \label{fig:bow_example}
  \end{figure}

@@

  \subsection{From Images to Text Documents}	
The original pipeline consists of the following 
steps: (1) region detection, (2) feature extraction, (3) vector quantization 
and (4) BoW histogram formation as in figure \ref{fig:bow_pipeline}. Tsai 
summarises the most common methods used recently \cite{tsai2012bag}.

    \begin{figure}[!ht]
    \centering
    \includegraphics[width=0.75\textwidth]{../figs/tsai2012}
    \caption{Bag of Words pipeline. The figure comes from \cite{tsai2012bag}}
    \label{fig:bow_pipeline}
    \end{figure}

@@

    A number of more advanced recipes for salient region localisation have been 
developed and implemented. These include for example SIFT
and ISS, with the vast majority 
implemented for the 2D domain and only a few generalised to 3D. A comparative 
study of detectors available in the PointCloud Library (PCL) can be found in 
\cite{pcl_keypoint_comparision, 3d_keypoint_eval}. All of the above, called 
sparse feature detectors, resort to selection of maxima in specific state 
spaces. An entirely different scheme is to use a dense feature detector, that 
is to take a uniformly sampled grid of points. Dense detectors are advantageous 
in that they sample slow changing regions in terms of image derivatives. 
Examples of such regions are a clear sky or a calm ocean. It was
showed that dense detectors generally outperform sparse ones 
\cite{tsai2012bag}.

  \subsubsection{Feature Extraction}
    Suppose only keypoint's coordinates were given. The smallest affine 
transformation applied to the image would render the keypoints useless. Feature 
extraction is about describing a region in a way that is invariant to affine 
transformations, light intensity changes or colour saturation. Achieving all 
these properties simultaneously can be hard but methods that achieve some of 
them do exist. Keypoint description takes form of coordinates in a 
high-dimensional space. One of the most precise and repeatable algorithm is 
SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 
128-dimensional vector of floating point values. It is the most often extracted 
descriptor in BoW pipelines. Other methods include various colour descriptors, 
binary descriptors such as 512-dimensional GIST. There are 
techniques designed for 3D exclusively. Among them are Persistent Point Feature 
Histogram (PFH) , its faster alternative FPFH 
 and PFHRGB, which takes into account colour information, 
all implemented in PCL.

@@

To build a dictionary we have to describe the training set (the previous steps) 
and find patterns in the descriptor space. A codebook can then contain 
centroids (means or medoids) of these patterns. The parsing phase is about 
assigning a descriptor to one of the codebook elements (e.g. 1-nearest 
neighbour matching with a centroid). A parsed descriptor is called a 
\textit{visual word}. Finally, we build a histogram by counting how many times 
each visual word occured for an image. The result is a fixed-size histogram, 
whose each bin tells us how many times a visual word occured.

    The kMeans is the single most popular vector quantization algorithm used in 
the BoW pipelines \cite{tsai2012bag}. Developed in 1950's, it is well known and 
easy to implement. kMeans divides all data into a predefined amount of clusters 
and computes their centroids. Many modifications and alternative versions 
emerged \cite{kmeans_jain2010data}. Some of them are: faster than the original 
\emph{approximate kmeans}, \emph{hierarchical kmeans}, which automatically 
chooses the final number of clusters and a \emph{soft kmeans} --- a variation 
that allows fuzzy alignment. Fuzzy alignment means that each point can belong 
to several clusters with different weights (e.g. proportional to the inverse 
square of a distance to a centroid).

    If computational cost is of no concern or if required precision is of the 
utmost priority we can use Gaussian Mixture Model (GMM). It partitions data 
into a set of clusters, finds their means and covariances. In the parsing step 
it computes a probability distribution of a descriptor over all the clusters. 
Fuzzy aligned histogram is built by taking the probability distribution as 
weights. GMM can be thought of as a generalisation of the kMeans algorithm. One 
drawback is massive computational cost in comparison with still expensive 
kMeans. Recently, Perronnin \textit{et al} proposed GMM based fisher-kernel 
vector quantization step yielding superior results \cite{fisher1, fisher2}.

    After choosing a vector quantization algorithm a codebook size have to be 
determined. With a standard kMeans implementation the dictionary size defines 
how much information is retained in a BoW histogram. A big trainset containing 
many classes with large intra- and inter-class variance is likely to 
require a huge codebook. On the other hand, too big a dictionary might 
introduce quantization artefacts. We have to bear in mind that the vector 
quantization step is the most time-consuming part of the BoW pipeline. The 
highest computational complexity is associated with the codebook construction 
step and it is $O(n^3)$, where $n$ is the codebook size.

@@

    To classify means to, given a set of categories, produce a category label 
for a given set of features. Image classification fits this 
description perfectly. The only difficulty lies in complexity of a raw image. 
Suppose we have a $320$ by $240$ three channel colour image. This amounts to a 
total of $230400$ dimensions, far too many to feed into any classifier 
directly. Fortunately, a Bag of Words histogram is a compact and discriminative 
intermediate representation that solves this issue. It can be used with any 
classifier. Classifiers can be two-class classifiers (e.g. binary) or 
multi-class classifiers. The latter are often a combination of several binary 
classifiers. The most popular classifiers are: k-Nearest Neighbours, Logistic 
Regression, Softmax Regression, Na\`ive Bayes and Support Vector Machine.

@@

  \subsection{Berkeley 3D Object Dataset}
    \begin{figure}[!ht]
    \centering	
    \includegraphics[width=.75\textwidth]{../figs/b3do_objects}
    \caption{B3DO objects: a) cups b) bottles c) books d) keyboards}
    \label{fig:b3do_objects}
    \end{figure}

    The Berkeley 3D Object dataset, shown in figure \ref{fig:b3do_objects}, was 
specifically designed for the purpose of object detection and classification. 
It consists of cluttered images in indoor environment, with many labelled and 
random objects per image. The images are often poorly lit and objects are 
partially occluded. There are around 50 classes with more than 20 examples in 
each of them. RGBD data are provided as pairs of colour images (8 bit RGB jpeg 
files) and depth maps (16 bit png files). Images are densely labelled --- for 
every pair of images there is an xml file with annotations. Each annotation 
states a category of an object as well as it's bounding box location. Since 
neither image segmentation nor object detection is addressed in this paper we 
had to extract objects from the images before the dataset could be used.

  \subsection{The University of Tokyo Dataset}	
    \begin{figure}[!ht]
    \centering
    \includegraphics[width=1\textwidth]{../figs/tokyo_horizontal}
    \caption{The Tokyo dataset: a) basket b) bicycle c) box d) bucket e) cart 
f) freezer g) notebook h) printer i) scanner j) scene}
    \label{fig:tokyo}
    \end{figure}

    The Tokyo dataset, shown in figure \ref{fig:tokyo}, is comprised of high 
quality images of single labelled objects in different settings, with changing 
viewpoint, object orientation and texture. Additional unlabelled objects can be 
present, however, and should not be taken into account. No bounding boxes are 
given, which renders the classification task similar to the scene 
classification problem. The dataset is presented as colour images (8 bit RGB 
jpeg files) and depth maps (csv files with XYZ coordinates of each pixel). This 
dataset is aimed at the task of object classification in casual images.

@@

    \begin{figure}[!ht]
    \centering	
    \includegraphics[width=.75\textwidth]{../figs/clustering_centroids_b3do}
    \caption{Influence of dictionary size on the overall accuracy. B3DO dataset 
with ISS keypoint detector and FPFH feature descriptor}
    \label{fig:cluster_b3do}
    \end{figure}

@@

    \begin{table}[!ht]
    \centering
    \caption{Results on the B3DO dataset with ISS keypoint detector, FPFH 
features and a dictionary of 1500 words; \textbf{Average accuracy: 65.22\%}}
    \includegraphics[width=0.9\textwidth]{../figs/b3do_conf_matrix}	
    \label{tab:b3do_conf_matrix}
    \end{table}

@@

  \subsection{Tokyo}
    \begin{table}[!ht]
    \centering
    \caption{Results on the Tokyo dataset with ISS keypoint detector, PFH 
features and a dictionary of 3000 words; \textbf{Average accuracy: 62.30\%}}
    \includegraphics[width=1\textwidth]{../figs/tokyo_conf_matrix}	
    \label{tab:tokyo_conf_matrix}
    \end{table}

@@

  In this paper the issue of three dimensional point cloud classification with 
a bag of words-based approach was tackled. Two scientific databases suitable 
for the task of point cloud classification were identified and adapted for the 
purpose of this work. Presence of such datasets is vital --- they act as 
benchmarks and allow objective comparison of different solutions of the 
problem. The following problems emerged: (1)~Point cloud processing is 
computationally expensive and inefficient. Due to the lack of structure in 
point clouds (\textit{i.e.} no predefined spatial relationships between points) 
additional calculations have to be performed (\textit{e.g.} finding nearest 
neighbours requires at least several distance computations). (2)~There are very 
few and small databases available. Machine learning algorithms have ravenous 
appetite for labelled data. As the newest findings in the state-of-the-art 2D 
object classification suggest, an order of tens of thousands of samples is a 
prerequisite for any satisfactory results. This work had to make do with less 
than 300 objects in the train sets. (3)~There is very little interest in the 
scientific community for this particular task. The majority of research is 
conducted on retrieval of CAD-like objects from mesh databases.

  The achieved results the B3DO and Tokyo datasets are respectively $5.21$ and 
$6.23$ times higher than random rates ($12.5\%$ and $10\%$). This shows that 
classification of point clouds is feasible and can be performed in what is 
close to real time on a mainstream notebook. The accuracy can be further 
enhanced by generalising 2D algorithms to the 3D domain, while the 
computational time can be reduced by using a GPU for the most computationally 
expensive parts of the Bag of Words framework. In conclusion, the depth data 
does not account for a straightforward improvement in classification accuracy. 
In fact, it is easier to obtain better results on 2D RGB images, because the 
best algorithms are not converted to the 3D domain and there is substantially 
more data available for the 2D.

>>>>

\title{Bag of Words Semantic Object Classification Based on RGBD Images}
\author{Adam Kosiorek}
\maketitle
e-mail: kosiorek.adam@gmail.com
Affiliation: Institute of Automation and Robotics, The Faculty of Mechatronics 
of Warsaw University of Technology \\
Supervisor: Prof. dr hab. Barbara Siemiątkowska \\
Defense date: 11.02.2014 \\
Keywords: bag-of-visual-words, object-classification, rgbd, point-cloud, datasets

@@

This paper reviews a Bag of Words (BoW) semantic object classification framework 
based on RGBD data registered by the Microsoft Kinect. We describe a BoW 
processing pipeline (region detection, feature extraction and vector 
quantization) and evaluate the most popular algorithms in terms of object 
classification performance. Salient region detection is done by SIFT, ISS3D and 
Uniform Sampling. Description is performed by FPFH, PFH and PFHRGB. Vector 
quantization was carried out by KMeans.  All-vs-All-trained SVM with the RBF 
kernel was used as a classifier. We review available point cloud databases in 
terms of usability in scene classification studies. We achieved the accuracy of 
65.22\% on 8 categories of the Berkeley 3D Object Dataset and 62.3\% on a 10 
category dataset compiled by Zhang.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%
%	Po Polsku !!!
%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Semantyczna klasyfikacja obiektów na obrazach RGBD metodą Bag of Words}
Słowa kluczowe: bag-of-visual-words, object-classification, rgbd, chmura-punktów, zbiór-danych
\abstract{

W pracy przedstawiono metodę klasyfikacji metodą Bag of Words (Bow) obrazów RGBD zarejstrowanych kamerą 
Microsoft Kinect. Omówiono proces przetwarzania BoW 
(detekcja i opis punktów, kwantyzacja i klasyfikacja) oraz zbadano 
najpopularniejsze algorytmy ze względu na efektywność klasyfikacji obiektów. 
Detekcja odbywa się za pomocą algorytmów SIFT, ISS3D i Uniform Sampling; do 
opisu punktów wykorzystano FPFH, PFH i PFHRGB; kwantyzację wykonano za pomocą 
algorytmu KMeans; jako klasyfikator użyto SVM z jądrem RBF trenowany metodą 
All-vs-All. Dokonano przeględu baz danych chmur punktów pod względem 
przydatności w badaniach klasyfikacji obiektów. Uzyskano skuteczność 
klasyfikacji 65.22\% na zbiorze 8 kategorii z Berkely 3D Object Dataset oraz 
62.3\% na zabiorze 10 kategorii udostępnionych przez Zhanga.

Podpisy pod rysunkami w kolejności występowania:
1. histogram w reprezentacji Bag of Words
2. Proces przetwarzania Bag of Words. Grafika pochodzi z [9]
3. obiekty ze zbioru B3D) a) kubki b) butelki c) książki d) klawiatury
4. Zbiór danych Tokyo a) koszyk b) rower c) pudełko d) koszyk e) wózek f) zamrażarka g) notebook h) drukarka i) skaner j) scena
5. Wpływ rozmiaru słownika na trafność klasyfikacji. Zbiór B3DO z detektorem ISS i deskryptorem FPFH

Opisy tabel:
1. Najwyższa trafność klasyfikacji z deskryptorami FPFH, PFH, PFHRGB na zbiorze B3DO
2. Wyniki na zbiorze B3DO z detektorem ISS, deskryptorem FPFH i słownikiem zawierającym 1500 słów
3. Wyniki na zbiorze Tokyo z detektorem ISS i deskryptorem PFH i słownikiem zawierającym 3000 słów

%%%%%%%%%%%%%%
%	Treść
%%%%%%%%%%%%%%

@@

  It has never been easier to capture visual information. Popular storage 
services grow lighting fast due to terabytes of photographs and movies uploaded 
every day. The growth is so fast that hand-tagging and description, the 
traditional means of annotation, cease to suffice. They are ambiguous, 
emotional and rarely optimal. With no better solution at hand databases are 
becoming increasingly harder to browse. Similar issues occur in mobile 
robotics, where a robot has to understand it's environment if it is to interact 
with it. Both problems can be cast as a scene or object classification problem, 
one of the most popular computer vision issues in the last decade.

  This paper reviews the Bag of Words (or BoW for short) approach to image 
classification, evaluates available point cloud databases in terms of usability 
in scene classification studies and applies BoW model in the 3D domain. The 
rest of this work is organized as follows: Section 2 describes the Bag of Words 
technique in detail. Construction of an image model is discussed and the most 
popular algorithms are outlined and the problem of classification is 
introduced. Section 3 describes two renown RGBD image datasets suitable for 
evaluation of RGBD data-based object classifiers. Section 4 addresses the 
topics of experimental setup, conducted experiments and provides the results. 
Finally, Section 5 concludes this work.

@@

  The Bag of Words model originated from the natural language processing domain 
\cite{tsai2012bag}. It 
assumes that word occurrence in a text document carries semantic information. 
It is an intermediate representation (in a form of a histogram)
used in natural language processing, information retrieval and general data 
mining.

  \begin{figure}[!ht]
  \centering
  \includegraphics[width=0.7\textwidth]{figs/bow_example}
  \caption{Bag of Words histogram}
  \label{fig:bow_example}
  \end{figure}

@@

  \subsection{From Images to Text Documents}	
The original pipeline consists of the following 
steps: (1) region detection, (2) feature extraction, (3) vector quantization 
and (4) BoW histogram formation as in figure \ref{fig:bow_pipeline}. Tsai 
summarizes the most common methods used recently \cite{tsai2012bag}.

    \begin{figure}[!ht]
    \centering
    \includegraphics[width=0.75\textwidth]{figs/tsai2012}
    \caption{Bag of Words pipeline. The figure comes from \cite{tsai2012bag}}
    \label{fig:bow_pipeline}
    \end{figure}

@@

    A number of more advanced recipes for salient region localization have been 
developed and implemented. These include for example SIFT
and ISS, with the vast majority 
implemented for the 2D domain and only a few generalized to 3D. A comparative 
study of detectors available in the PointCloud Library (PCL) can be found in 
\cite{pcl_keypoint_comparision, 3d_keypoint_eval}. All of the above, called 
sparse feature detectors, resort to selection of maxima in specific state 
spaces. An entirely different scheme is to use a dense feature detector, that 
is to take a uniformly sampled grid of points. Dense detectors are advantageous 
in that they sample slow changing regions in terms of image derivatives. 
Examples of such regions are a clear sky or a calm ocean. It was
showed that dense detectors generally outperform sparse ones 
\cite{tsai2012bag}.

  \subsubsection{Feature Extraction}
    Suppose only keypoint's coordinates were given. The smallest affine 
transformation applied to the image would render the keypoints useless. Feature 
extraction is about describing a region in a way that is invariant to affine 
transformations, light intensity changes or color saturation. Achieving all 
these properties simultaneously can be hard but methods that achieve some of 
them do exist. Keypoint description takes form of coordinates in a 
high-dimensional space. One of the most precise and repeatable algorithm is 
SIFT \cite{sift_features}, which is a 3D histogram of gradients structured as a 
128-dimensional vector of floating point values. It is the most often extracted 
descriptor in BoW pipelines. Other methods include various color descriptors, 
binary descriptors such as 512-dimensional GIST. There are 
techniques designed for 3D exclusively. Among them are Persistent Point Feature 
Histogram (PFH) , its faster alternative FPFH 
 and PFHRGB, which takes into account color information, 
all implemented in PCL.

@@

To build a dictionary we have to describe the training set (the previous steps) 
and find patterns in the descriptor space. A codebook can then contain 
centroids (means or medoids) of these patterns. The parsing phase is about 
assigning a descriptor to one of the codebook elements (e.g. 1-nearest 
neighbor matching with a centroid). A parsed descriptor is called a 
\textit{visual word}. Finally, we build a histogram by counting how many times 
each visual word occurred for an image. The result is a fixed-size histogram, 
whose each bin tells us how many times a visual word occurred.

    The KMeans is the single most popular vector quantization algorithm used in 
the BoW pipelines \cite{tsai2012bag}. Developed in 1950's, it is well known and 
easy to implement. KMeans divides all data into a predefined amount of clusters 
and computes their centroids. Many modifications and alternative versions 
emerged \cite{kmeans_jain2010data}. Some of them are: faster than the original 
\emph{approximate KMeans}, \emph{hierarchical KMeans}, which automatically 
chooses the final number of clusters and a \emph{soft KMeans} --- a variation 
that allows fuzzy alignment. Fuzzy alignment means that each point can belong 
to several clusters with different weights (e.g. proportional to the inverse 
square of a distance to a centroid).

    If computational cost is of no concern or if required precision is of the 
utmost priority we can use Gaussian Mixture Model (GMM). It partitions data 
into a set of clusters, finds their means and covariances. In the parsing step 
it computes a probability distribution of a descriptor over all the clusters. 
Fuzzy aligned histogram is built by taking the probability distribution as 
weights. GMM can be thought of as a generalization of the KMeans algorithm. One 
drawback is massive computational cost in comparison with still expensive 
KMeans. Recently, Perronnin \textit{et al} proposed GMM based fisher-kernel 
vector quantization step yielding superior results \cite{fisher1, fisher2}.

    After choosing a vector quantization algorithm a codebook size have to be 
determined. With a standard KMeans implementation the dictionary size defines 
how much information is retained in a BoW histogram. A big trainset containing 
many classes with large intra- and inter-class variance is likely to 
require a huge codebook. On the other hand, too big a dictionary might 
introduce quantization artefacts. We have to bear in mind that the vector 
quantization step is the most time-consuming part of the BoW pipeline. The 
highest computational complexity is associated with the codebook construction 
step and it is $O(n^3)$, where $n$ is the codebook size.

@@

    To classify means to, given a set of categories, produce a category label 
for a given set of features. Image classification fits this 
description perfectly. The only difficulty lies in complexity of a raw image. 
Suppose we have a $320$ by $240$ three channel color image. This amounts to a 
total of $230400$ dimensions, far too many to feed into any classifier 
directly. Fortunately, a Bag of Words histogram is a compact and discriminative 
intermediate representation that solves this issue. It can be used with any 
classifier. Classifiers can be two-class classifiers (e.g. binary) or 
multi-class classifiers. The latter are often a combination of several binary 
classifiers. The most popular classifiers are: k-Nearest Neighbors, Logistic 
Regression, Softmax Regression, Na\`ive Bayes and Support Vector Machine.

@@

  \subsection{Berkeley 3D Object Dataset}
    \begin{figure}[!ht]
    \centering	
    \includegraphics[width=.75\textwidth]{figs/b3do_objects}
    \caption{B3DO objects: a) cups b) bottles c) books d) keyboards}
    \label{fig:b3do_objects}
    \end{figure}

    The Berkeley 3D Object dataset, shown in figure \ref{fig:b3do_objects}, was 
specifically designed for the purpose of object detection and classification. 
It consists of cluttered images in indoor environment, with many labeled and 
random objects per image. The images are often poorly lit and objects are 
partially occluded. There are around 50 classes with more than 20 examples in 
each of them. RGBD data are provided as pairs of color images (8 bit RGB jpeg 
files) and depth maps (16 bit png files). Images are densely labeled --- for 
every pair of images there is an xml file with annotations. Each annotation 
states a category of an object as well as it's bounding box location. Since 
neither image segmentation nor object detection is addressed in this paper we 
had to extract objects from the images before the dataset could be used.

  \subsection{The University of Tokyo Dataset}	
    \begin{figure}[!ht]
    \centering
    \includegraphics[width=1\textwidth]{figs/tokyo_horizontal}
    \caption{The Tokyo dataset: a) basket b) bicycle c) box d) bucket e) cart 
f) freezer g) notebook h) printer i) scanner j) scene}
    \label{fig:tokyo}
    \end{figure}

    The Tokyo dataset, shown in figure \ref{fig:tokyo}, is comprised of high 
quality images of single labeled objects in different settings, with changing 
viewpoint, object orientation and texture. Additional unlabeled objects can be 
present, however, and should not be taken into account. No bounding boxes are 
given, which renders the classification task similar to the scene 
classification problem. The dataset is presented as color images (8 bit RGB 
jpeg files) and depth maps (csv files with XYZ coordinates of each pixel). This 
dataset is aimed at the task of object classification in casual images.

@@

    \begin{figure}[!ht]
    \centering	
    \includegraphics[width=.75\textwidth]{figs/clustering_centroids_b3do}
    \caption{Influence of dictionary size on the overall accuracy. B3DO dataset 
with ISS keypoint detector and FPFH feature descriptor}
    \label{fig:cluster_b3do}
    \end{figure}

@@

    \begin{table}[!ht]
    \centering
    \caption{Results on the B3DO dataset with ISS keypoint detector, FPFH 
features and a dictionary of 1500 words; \textbf{Average accuracy: 65.22\%}}
    \includegraphics[width=0.9\textwidth]{figs/b3do_conf_matrix}	
    \label{tab:b3do_conf_matrix}
    \end{table}

@@

  \subsection{Tokyo}
    \begin{table}[!ht]
    \centering
    \caption{Results on the Tokyo dataset with ISS keypoint detector, PFH 
features and a dictionary of 3000 words; \textbf{Average accuracy: 62.30\%}}
    \includegraphics[width=1\textwidth]{figs/tokyo_conf_matrix}	
    \label{tab:tokyo_conf_matrix}
    \end{table}

@@

  In this paper the issue of three dimensional point cloud classification with 
a bag of words-based approach was tackled. Two scientific databases suitable 
for the task of point cloud classification were identified and adapted for the 
purpose of this work. Presence of such datasets is vital --- they act as 
benchmarks and allow objective comparison of different solutions of the 
problem. The following problems emerged: (1)~Point cloud processing is 
computationally expensive and inefficient. Due to the lack of structure in 
point clouds (\textit{i.e.} no predefined spatial relationships between points) 
additional calculations have to be performed (\textit{e.g.} finding nearest 
neighbors requires at least several distance computations). (2)~There are very 
few and small databases available. Machine learning algorithms have ravenous 
appetite for labeled data. As the newest findings in the state-of-the-art 2D 
object classification suggest, an order of tens of thousands of samples is a 
prerequisite for any satisfactory results. This work had to make do with less 
than 300 objects in the train sets. (3)~There is very little interest in the 
scientific community for this particular task. The majority of research is 
conducted on retrieval of CAD-like objects from mesh databases.

  The achieved results the B3DO and Tokyo datasets are respectively $5.21$ and 
$6.23$ times higher than random rates ($12.5\%$ and $10\%$). This shows that 
classification of point clouds is feasible and can be performed in what is 
close to real time on a mainstream notebook. The accuracy can be further 
enhanced by generalizing 2D algorithms to the 3D domain, while the 
computational time can be reduced by using a GPU for the most computationally 
expensive parts of the Bag of Words framework. In conclusion, the depth data 
does not account for a straightforward improvement in classification accuracy. 
In fact, it is easier to obtain better results on 2D RGB images, because the 
best algorithms are not converted to the 3D domain and there is substantially 
more data available for the 2D.
